A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-03-25 paper.tex
Better overview of bounded synth by removing discussion of quantifiers

Realisability for a bounded safety game can be solved by checking a quantified
formula consisting of the game's transition relation \emph{unrolled} to the
length of the bounded game. The quantifiers represent the two players of the
game: universal for the controller and existential for the environment.  Just
as the players alternate in choosing actions every round of the game, the
formula is prefixed by a quantifier alternation over $U$ and $C$ for each
round.

Instead of solving both quantifiers at once, two competing solvers are used to
solve abstractions of the game. Each solver builds a candidate strategy for its
corresponding player and uses its opponent's candidate to restrict its own
search. The solvers communicate to refine one another's candidate strategies by
finding counterexamples. Eventually one solver will be unable to find a
counterexample, indicating that its opponent's candidate strategy is a winning
strategy for the game.

@@

\begin{algorithm}[t]
    \begin{algorithmic}[1]
        \Require $s(X_T) \land $ \Call{treeFormula}{$k, T$} $\equiv \bot$
        \Require \emph{Must-invariant} holds
        \Ensure \emph{Must-invariant} holds
        \Ensure $s(X_T) \land B^M \not\equiv \bot$
        \Comment $s$ will be added to $B^M$
        \Function{learn}{$s, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{treeFormula}{$k, T_1$}, \Call{treeFormula}{$k, T_2$}}
            \State $B^M \gets B^M \lor \mathcal{I}$
            \State \Call{learn}{$s, T_1$}
        \EndFunction
        \algstore{learn}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{learn}
        \Require $s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T$} $\equiv \bot$
        \Require \emph{May-invariant} holds
        \Ensure \emph{May-invariant} holds
        \Ensure $s(X_T) \land B^m[$\Call{height}{$k, T$}$] \equiv \bot$
        \Comment{$s$ will be removed from $B^m[$\Call{height}{$k, T$}$]$}
        \Function{\textoverline{learn}}{$s, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T_1$}, \Call{\textoverline{treeFormula}}{$k, T_2$}}
            \For{$i = 1$ to \Call{height}{$k, n$}}
                \State $B^m[i] \gets B^m[i] \setminus \mathcal{I}$
            \EndFor
            \State \Call{\textoverline{learn}}{$s, T_1$}
        \EndFunction
    \end{algorithmic}
    \caption{Learning algorithms}
    \label{alg:learn}
\end{algorithm}

@@

\begin{algorithm}[h]
    \begin{algorithmic}[1]
        \Function{solveUnbounded}{$T$}
            \State $B^M \gets E$
            \State $B^m[0] \gets \top$
            \For{$k = 1 \dots$}
                \IIf{\Call{SAT}{$I \land B^M$}}
                    \Return \texttt{unrealisable} \Comment{Losing in the initial state}
                \EndIIf
                \IIf{$\exists i < k . \  B^m[i] \equiv B^m[i+1]$} \Comment{Reached fixed point}
                    \State \hspace{\algorithmicindent} \Return \texttt{realisable} 
                \EndIIf
                \State $B^m[k] \gets \top$
                \State \Call{checkBound}{$k$}
            \EndFor
        \EndFunction
        \algstore{u1}
    \end{algorithmic}

>>>>

The algorithm consists of two competing solvers, one for the system and one for
the environment, that solve abstractions of the safety game. Each solver builds
a candidate strategy for its corresponding player consisting of assignments to
$2^C$ and $2^U$ respectively.

The game abstraction is initially constructed from an opponent candidate
strategy. It is refined by counterexamples to the solver's own candidate
strategy that are formed when the opponent solves the abstract game
corresponding to that strategy. By abstracting the game in this way we
conceptually direct the solvers' search to consider opponent actions that may
lead to winning strategies.

@@

\begin{algorithm}[t]
    \begin{algorithmic}[1]
        \Require $s(X_T) \land $ \Call{treeFormula}{$k, T$} $\equiv \bot$
        \Require \emph{Must-invariant} holds
        \Ensure \emph{Must-invariant} holds
        \Ensure $s(X_T) \land B^M \not\equiv \bot$
        \Comment $s$ will be added to $B^M$
        \Function{learn}{$s, T$}
            \IIf{\Call{succ}{$T$} $= \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{treeFormula}{$k, T_1$}, \Call{treeFormula}{$k, T_2$}}
            \State $B^M \gets B^M \lor \mathcal{I}$
            \State \Call{learn}{$s, T_1$}
        \EndFunction
        \algstore{learn}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{learn}
        \Require $s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T$} $\equiv \bot$
        \Require \emph{May-invariant} holds
        \Ensure \emph{May-invariant} holds
        \Ensure $s(X_T) \land B^m[$\Call{height}{$k, T$}$] \equiv \bot$
        \Comment{$s$ will be removed from $B^m$}
        \Function{\textoverline{learn}}{$s, T$}
            \IIf{\Call{succ}{$T$} $= \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T_1$}, \Call{\textoverline{treeFormula}}{$k, T_2$}}
            \For{$i = 1$ to \Call{height}{$k, n$}}
                \State $B^m[i] \gets B^m[i] \setminus \mathcal{I}$
            \EndFor
            \State \Call{\textoverline{learn}}{$s, T_1$}
        \EndFunction
    \end{algorithmic}
    \caption{Learning algorithms}
    \label{alg:learn}
\end{algorithm}

@@

\begin{algorithm}[h]
    \begin{algorithmic}[1]
        \Function{solveUnbounded}{$T$}
            \State $B^M \gets E$
            \State $B^m[0] \gets E$
            \For{$k = 1 \dots$}
                \IIf{\Call{SAT}{$I \land B^M$}}
                    \Return \texttt{unrealisable} \Comment{Losing in the initial state}
                \EndIIf
                \IIf{$\exists i < k . \  B^m[i] \equiv B^m[i+1]$} \Comment{Reached fixed point}
                    \State \hspace{\algorithmicindent} \Return \texttt{realisable} 
                \EndIIf
                \State $B^m[k] \gets E$
                \State \Call{checkBound}{$k$}
            \EndFor
        \EndFunction
        \algstore{u1}
    \end{algorithmic}
